---
title: "Final Exam (Makeup)"
author: "Prof. Mahdavi"
date: "Winter 2026"
output: html_document
---

TIME LIMITS:
YOU HAVE 24 HOURS TO COMPLETE THIS EXAM.

YOU MUST SUBMIT YOUR CODE AND HTML FILE VIA EMAIL BY 12:00PM FRIDAY MARCH 20, 2026.

**If you cannot knit your code for any reason, you should still submit your code (.Rmd) file via email. Submissions without a knitted HTML file will incur a 2-point penalty.**

RULES:
YOU MAY USE THE INTERNET, AND ANY MATERIAL FROM THE COURSE.

YOU MAY NOT USE GENERATIVE AI. IF WE SUSPECT USAGE OF AI TO COMPLETE THIS EXAM, 
IT WILL BE AN AUTOMATIC FAILURE OF THE EXAM AND THE COURSE. PLEASE ADHERE TO THIS RULE!!!!

NO EMAIL, PHONE, OR COMMUNICATION WITH ANYONE ABOUT THE FINAL DURING THE 3-HOUR WINDOW.
FAILURE TO FOLLOW THIS RULE LEADS TO A ZERO IN THE COURSE.

PLEASE WRITE YOUR CODE AND YOUR TEXT ANSWER BELOW EACH QUESTION IN THE CODE CHUNK AND THE SPACE MARKED *Answer:* 

### Part 1. Short answer questions (30 pts total)

Q1. You toss a fair coin (i.e., one that gives you heads about half the time) 1000 times and record the outcome for each toss. If you were to graph a histogram or density plot of all the outcomes, would you roughly see a Bernoulli distribution? Explain why or why not in no more than two sentences, being sure to clearly define this distribution and the broader type of distribution it represents. **(5 pts. total)**

*Answer:* 


Q2. True or False: In the model $Y = B_0 + B_1 X_1 + B_2 X_2 + \epsilon$, our estimate of $B_1$ will be biased if we omit a variable $X_3$ that is correlated with $X_1$ and $X_2$ but does not have a relationship with $Y$. Please answer whether this statement is true or false, and explain why in no more than two sentences. **(5 pts. total)**

*Answer:* 


Q3. We read an article this quarter by Regina Bateson. (3-4 sentences, 100 words max) **(6 pts. total)**
What is her research question and what is her dependent variable? *(1 pt.)*
What does she find? *(1 pt.)*
Are her results robust to alternative model specifications? How does she test this? *(1 pt.)*
Does her research design allow for causally identifying an effect? *(2 pts.)*

*Answer:* 

Q5. Your friend asks you to explain what a random variable is. In your own words, provide a definition. What are its key components? **(4 pts.)**

*Answer:* 

Q6. Your friend then says they have heard of linear regression before, but they don’t know how it works. Explain in simple language what the regression is doing to estimate a relationship between two variables. Come up with a specific political science example to help your friend understand. **(6 pts.)**

*Answer:* 

Q7. We also read an article by Gerber, Green, and Larimer (2008), where the authors ran an experiment with four treatment options, each representing a type of pressure, to show that social norms affect political participation. List each of these treatments and describe how the regression results for these treatments connect with the findings of the study. **(4 pts.)**

*Answer:*


### Part 2. Do property rights undermine development? (70 pts. total)

Political economy theory views property rights as the foundation of economic and political development. But do property rights that are too strong encourage opportunistic behavior? To answer this question, political scientist Alisha Holland designed a survey experiment in Colombia. Her study randomly assigned individuals to treatment groups with different property rights frames and asked them about their preferences for highway and construction projects funded by the state. The key treatment is whether or not a respondent saw the following message: "The state protects the right to private property because property owners invest in their homes and neighborhoods." You will be replicating one aspect of her study, which shows how property rights affect attitudes towards compensation for state-led highway projects that go through an individual's land.

Load the dataset, `survey.RData`. Here is a brief description of the variables:

* responseid - unique survey respondent ID number
* proprights - dummy variable for whether individual received property rights treatment (0 = no, 1 = yes)
* compensation - amount in thousands of pesos that respondent desires from government in compensation for highway going through their property
* female - dummy variable for whether respondent is female or not (0 = no, 1 = yes)
* age - age of respondent in five ordered groups (1 = 18-24, ..., 5 = 55 or older)
* education - dummy variable for whether respondent finished high school or not (0 = no, 1 = yes)
* wealth - wealth index of respondent in deciles (1 = poorest, ..., 10 = richest)

Q8. Set your working directory and load the dataset `survey.RData`. **(4 pts. total)**

```{r}

```


Q9. It's always good practice to begin by looking at and summarizing your data. **(6 pts. total)** 
What is your sample size? *(1 pt.)*
What is the mean value of `compensation`? *(1 pt.)*
What is the variance of this variable? *(1 pt.)*
What is the standard deviations of this variable? *(1 pt.)*
What is the range of this variable? *(1 pt.)*
What is the unit of analysis? *(1 pt.)*

```{r}

```

*Answer:* 


Q10. Now let's look at the distribution of `proprights`. **(8 pts. total)**
Plot the histogram of this variable. *(2 pt.)*
Interpret the histogram plot - how could you describe this distribution? *(3 pt.)*
How would you describe the *distribution of the mean* of this distribution? *(3 pt.)*

```{r}

```

*Answer:* 

Q11. Now let's examine the association between `proprights` and `compensation`. **(8 pts. total)**
What is the covariance of `proprights` and `compensation`? Give the estimate and describe it in your own words. *(4 pt.)*
What is the correlation? Give the estimate and describe it in your own words. *(4 pt.)*

```{r}

```


Q12. You want to know whether the property rights treatment affects compensation, controlling for age, gender, and wealth. **(8 pts. total)**
In this case, what is your main independent variable? *(2 pt.)* 
What is your dependent variable? *(2 pt.)*
State your null hypothesis. *(2 pt.)* 
State your alternative hypothesis. *(2 pt.)*

*Answer:* 


Q13. Plot the relationship between `wealth` (X) and `compensation` (Y) using a scatterplot. For full credit, you must label all axes in your graph, including the title. **(4 pts. total)**

```{r}

```



Q14. Run a regression that shows the relationship between the property rights treatment (X) and compensation (Y), controlling for age, female, and wealth. **(7 pts. total)**
Show the output of the regression. *(2 pt.)*
Substantively interpret the coefficient that shows the relationship between property rights and compensation. *(5 pt.)*

```{r}

```

*Answer:* 


Q15. Interpret the p-value from the regression above for the property rights treatment. Next interpret what a p-value generally means in practice, using terms used in lecture -- not using terms found online! **(4 pts. total)**

*Answer:*


Q16. Intepret your findings in terms of your hypotheses. Did you find evidence for or against the null hypothesis? **(3 pts. total)**

*Answer:* 

Q17. Construct the 95% confidence interval for the  property rights treatment variable based on the results of the regression above. Does the confidence interval include 0? What does this tell you about whether or not we can reject the null hypothesis? **(4 pts. total)**

*Answer:* 



Q18. A critic challenges the findings of the study by claiming that the author did not account for potential imbalance between people who went completed high school and those who did not. To evaluate their criticism, add the education dummy variable as an additional control to your model from Q14 and run a new regression. **(5 pts. total)**
Show the output of this regression. *(1 pt.)*
Interpret the coefficient on the education dummy variable and evaluate its statistical significance. *(2 pt.)*
Are the model results from Q14 robust to this change? *(2 pt.)*

```{r}

```

*Answer:* 


Q19. Can you say that we have identified a causal relationship between property rights and compensation? Why or why not? **(4 pts. total)**

*Answer:* 


Q20. If you had infinite financial resources, what kind of research design would you construct to better assess the causal relationship between property rights and compensation? **(5 pts. total)**

*Answer:* 


## Ungraded Question (no points, but must be answered!).

Please confirm here that you did NOT use AI to complete this exam. 

*Answer:*
