---
title: "POLS 15 -- Section 6"
author: "Hunter C. Smith"
date: "`r Sys.Date()`"
output: html_document
---

### DIY Setup

As always, clear your environment, set your working directory, and load in today's data file, `wb.RData`.

```{r}
# Type your code here!
```

Here's a description of the variables we'll be using today:

Variable Name            | Description      
------------------------ | -----------------------------------------      
`GDPgrowth`              | Growth rate for gross domestic product, measured as a percentage.      
`CO2emissionscap`        | Carbon dioxide emissions per capita, measured as metric tons.  
`Healthexpendituretotal` | Health expenditures as a percentage of total GDP. 
`Mortalityrate`          | Number of deaths per 1,000 adults. 

### Hypothesis Testing

Today, we'll be covering various methods of hypothesis testing -- essentially, how we determine whether our estimates are considered statistically significant. Some of this has been briefly covered in previous sections and some will be new.

Let's remind ourselves of the null and alternate hypotheses we'll be evaluating -- what do each of these mean in words?

$H_0: \beta = 0$ 
$H_A: \beta \neq 0$

We'll begin with a review of $t$-statistics and $p$-values, both of which are reported in the regression summary table. Let's examine whether GDP growth rates predict carbon dioxide emissions with a simple bivariate regression.

```{r}
mod1 <- lm(CO2emissionscap ~ GDPgrowth, data = wb)
summary(mod1)
```

In the context of regression, $t$-statistics are calculated by dividing the coefficient estimate by its standard error: $t = \frac{\beta}{SE_{\beta}}$. The $t$-statistic associated with $\hat{\beta_1}$ is 0.006, and the associated $p$-value is 0.995. To assess statistical significance, we need to compare these to critical values. For this class, we choose critical values that are consistent with a 95% confidence level.

At 95% confidence, the critical $t$ value, $t^*$ is $\pm$1.96 -- $t$-statistics can be positive or negative, so we need to use the $\pm$ symbol to denote both possibilities. We can say that our estimate is statistically significant if $|t| > |t^*|$. In other words, if the absolute value of our $t$-statistic is larger than 1.96, our estimate is statistically significant. In this case, $t$ = 0.006, which is less than 1.96, so $\hat{\beta_1}$ is statistically insignificant. What decision do we make regarding the null hypothesis in this case?

Another way we can evaluate our hypotheses is by using the $p$-value. As you learned in lecture, the $p$-value tells us the probability of observing the estimate we calculated *if the null hypothesis is true*. In other words, if there is NOT a relationship between X and Y, what is the probability that we would observe the relationship we did? It quantifies the probability of a Type I error (a false positive). 

In this case, the $p$-value associated with $\hat{\beta_1}$ is 0.995. For $p$-values, our point of comparison at a 95% confidence level is 0.05 and is denoted as $\alpha$. An estimate is statistically significant if $p < \alpha$. Note the differences from evaluating the $t$-statistic: $p$-values are bounded between zero and one, so we don't need to worry about negatives or absolute values, and the direction of the comparison is reversed -- we want large $t$-statistics but small $p$-values. For the model above, $p$ = 0.995, which is larger than $\alpha$, so $\hat{\beta_1}$ is statistically insignificant. What decision do we make regarding the null hypothesis in this case?

So far, we've evaluated the significance of $\hat{\beta_1}$ in two ways. Confidence intervals represent a third way we can do this. A confidence interval tells us the range of possible $\beta_1$ values that are consistent with our estimate, $\hat{\beta_1}$. It has an upper and lower bound, and can be manually calculated using values from the regression table. As with $t^*$ and $\alpha$, the confidence interval changes depending on the level of confidence we use. Here is the general formula for a confidence interval:

$$CI = \hat{\beta_1} \pm t^* * SE_{\hat{\beta_1}}$$

Since we use a 95% confidence level, which corresponds to $t^* = \pm 1.96$, we can fill that value in:

$$CI = \hat{\beta_1} \pm 1.96 * SE_{\hat{\beta_1}}$$

Go ahead and calculate the upper and lower bound of the 95% confidence interval associated with $\hat{\beta_1}$ in the code chunk below.

```{r}
# Type your calculations here!
```

We determine statistical significance from confidence intervals based on whether zero is contained within the interval. If it is, this means that a null relationship is consistent with the non-null relationship we estimated, which means we cannot be confident a relationship truly exists. Based on the bounds you calculated, what do you conclude about significance?

One thing to note is that no matter which method you use, you should ALWAYS come to the same conclusion about significance for an estimate. This is a good check about your interpretations -- if you evaluate a $t$-statistic as indicating significance but evaluate a $p$-value as indicating insignificance, you've misinterpreted one of your values.

As a quick recap, here are the rules for determining significance for each method:
- $t$-statistic: $|t| > |t^*|$
- $p$-value: $p < \alpha$
- Confidence Interval: zero is not contained within the interval

### Visualizing Confidence Intervals

It's sometimes helpful for us to plot confidence intervals relative to our regression line. Let's examine a new question: does spending on health affect mortality rates? First, let's estimate a regression model and add the regression line to a scatterplot.

```{r}
mod2 <- lm(Mortalityrate ~ Healthexpendituretotal, data = wb)

plot(wb$Healthexpendituretotal, wb$Mortalityrate,
     xlab = "Health Expenditure",
     ylab = "Mortality Rate")

abline(mod2, col = "blue", lwd = 2)
```

The line indicates a negative relationship between the two -- as health expenditures increase, mortality rates decrease. Now, we're going to add both a 95% and a 99% confidence interval to our plot (you will NOT be asked to do this coding yourself).

```{r}
plot(wb$Healthexpendituretotal, wb$Mortalityrate,
     xlab = "Health Expenditure",
     ylab = "Mortality Rate")

abline(mod2, col = "blue", lwd = 2)

newx <- seq(min(wb$Healthexpendituretotal), max(wb$Healthexpendituretotal), length.out = 100)

CI95 <- predict(mod2, 
                newdata = data.frame(Healthexpendituretotal = newx),
                interval = "confidence",
                level = 0.95)
lines(newx, CI95[, 2], lty = 2, col = "darkgreen")
lines(newx, CI95[ ,3], lty = 2, col = "darkgreen")

CI99 <- predict(mod2, 
                newdata = data.frame(Healthexpendituretotal = newx),
                interval = "confidence", 
                level = 0.99)
lines(newx, CI99[, 2], lty = 3, col = "purple")
lines(newx, CI99[, 3], lty = 3, col = "purple")
```

Quick intuition check: why is the 99% confidence interval wider than the 95% confidence interval?

### Do It Yourself!

Now, evaluate this relationship using the three methods you've just learned!

Null Hypothesis: ???
Alternative Hypothesis: ???

Method 1: ???
Method 2: ???
Method 3: ???

### Knit to HTML

Always remember to knit your file when you finish anything!