---
title: "Problem Set 1 (Due 23 January 2026, 6:00 pm)"
author: "Prof. Paasha Mahdavi, PS 15, UCSB"
date: "Winter 2026"
output: html_document
---

# Introduction

Welcome to the first problem set. There is not a great deal of material here, but since this may be your first time using R and R Markdown, there are many potential pitfalls, so leave yourself plenty of time to complete it. The idea here is not that you can sit down and answer these questions straight away, but that you have a week to figure it out, and this is a key part of the learning process.

If you are looking at the HTML version of the problem set (`PSet1_W26.html`) that may have opened in your web browser, you are seeing the ouput produced by running the "script"" or code called in the file `PSet1_W26.Rmd`, also available on the course website. Go ahead and open the file called `PSet1_W26.Rmd`. If it does not open automatically within RStudio, you can open RStudio first and then use the File menu to open up `PSet1_W26.Rmd`. Once you open `PSet1_W26.Rmd`, you can continue reading the text easily in that file.

*It will be easiest for you to open the `.Rmd` file posted for each PSet, and start writing your solutions in by learning from the code you see in the questions.*

# Before you start: The R Markdown Introduction

The text, output and graphics in this section are provided as an example whenever you create a new R Markdown (`.Rmd`) file in RStudio. It's a good quick introduction so I replicate it here with minor modification. At this point, you may not understand all of the `R` code being used here, but the goal is to understand how the `.Rmd` file works and how it relates to the `.html` file that gets outputted.

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. (Here we have set the code to produce an HTML output, which is what you need to upload for this class).

When you click the **Knit** button a document will be generated that includes both the content as well as the output of any embedded `R` code chunks within the document. You can embed an `R` code chunk like this:

```{r}
summary(iris)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(x = iris$Sepal.Length, y = iris$Petal.Length)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the `R` code that generated the plot. In general, you will not include the `echo = FALSE`, because we will want to see your code.

# Submission Instructions

Please submit your problem set via Canvas. Submit both the `.Rmd` file, and the `.html` file it creates when you knit it. This assignment is due by 6pm on Friday, January 23, 2026. Please list any students you collaborated with. Please also disclose whether you employed generative AI to complete this assignment, and how you used it, in the space at the end of the problem set.

# Example Problem

Here is an example problem, with an example solution.

**Question 0.** In this question, we'll provide the answer for you, as an example. You need to be looking at the `.Rmd` file right now for this to make much sense.

Showing your code and the result, execute the code `getwd()`. Describe what this command does. You may want to execute the command directly in the console first (ask your TA if you don't know how to run a command--this is essential) to see what it does, but be sure to write it into your `.Rmd` file so that it runs when you click **knit**.

**Solution: Question 0**

```{r}
getwd()
```

This command, when executed (either in the console or through the .rmd file once you click \`\`knit''), tells the user what directory is set as the working directory. This is the directory where output will be saved, or where R will look first when searching for a file, for example a dataset. You need to always set your working directory first so R knows where to pull the data from.

------------------------------------------------------------------------

Make sure to try executing your `.Rmd` file now by clicking **knit**. Then take a look at the `.html` file that it created and see what you get.

Okay, your turn to answer the remaining questions!

*Part 1. Theory. Short answer questions (5 points total)*

Start by reading **Chapter 1** of the textbook, *Real Stats*. You can also review your lecture notes. You may **NOT** use generative AI to answer these questions.

Q1. A researcher observes that countries with democratic systems have weaker climate policies, compared to countries with non-democratic systems. She decides to publish a research article that says having this form of government causes countries to become less climate friendly. Would you like to be a co-author on this paper? Why or why not? (100 words max)

**ANSWER: **

Q2. Explain what this sentence means: "Experiments create exogeneity via randomization." (75 words max)

**ANSWER: **

Q3. What do we refer when we talk about the "internal validity" and "external validity" of experiments? (100 words max)

**ANSWER: **

Q4. You decide to run an experiment to see whether working in groups helps students learn. You randomly assign half of the class to form study groups to work on their problem sets, and to visit office hours every week. For the other half of the class, you assign them to work individually. At the end of the semester, you give the entire class a test. You find that the students in the first group did much better than those in the second group, who worked individually. (120 words max across all parts)

4a. What could you call each group? 

**ANSWER: **

4b. What is your independent variable and what is your dependent variable? 

**ANSWER: **

4c. Given this set up, list some factors you need to control for. 

**ANSWER: **

4d. Can you say that working in groups caused the students to do better on the test? Why or why not? Explain using the technical terms in the textbook. 

**ANSWER: **

4e. Can you say that this finding would also apply in other types of courses (for example, writing-intensive classes)? Why or why not? Explain using the technical terms in the textbook.

**ANSWER: **

Q5. Imagine you are looking at whether people with higher levels of education have higher incomes. List some of the factors that could lead to endogeneity. (50 words max)

**ANSWER: **

------------------------------------------------------------------------

*Part 2. Data Analysis (5 points total)*

The improvement in human rights in the second half of the 20th century is one of the most relevant global development trends. What is the relationship between economic development and human rights?

We will use a dataset from "Our World in Data (OWID)" (Saloni Dattani and Fiona Spooner and Hannah Ritchie and Max Roser, 2023) to explore this question.

Download the dataset, `humanrights.RData`, which you'll find on the course website and also on the online RStudio platform. You may want to put it in your working directory to make it easy to find (use `getwd()` to see what your current working directory is; you can use the Session menu in Rstudio or the `setwd()` command to change your working directory.)

Here is a brief description of the variables:

-   `country` - name of the country or territory
-   `humanrights` - average Human Rights Index, a measure ranging from 0 to 1 that reflects the level of human rights protections in a country
-   `gdppc` - average income of an individual in a country, measured as GDP per capita

*As will often be the case when using `R`, you will need to use the `$` operator to access these variables within the object. You can reference the section files if you need a reminder of how to do this. Remember, the end of each chapter in the textbook includes `R` code that can be helpful. We also posted `R` resources on Canvas.*

Q1. Load the data into `R`. The data are stored as an `.RData` file, so you can use the `load()` function to load it. What name is given to the dataset object in your environment?

**ANSWER: **

Q2. Check the dimensions of the data (i.e. the number of rows and columns). How many observations are there? How many variables are there?

**ANSWER: **

Q3. What are the names of the variables in the dataset?

**ANSWER: **

Q4. Calculate the average of `gdppc`, which is average income per capita for each country, across all points in the sample. Do you think this average is large or small? Explain your answer. What is the minimum and the maximum `gdppc`? Make sure to substantively explain what each of these values are communicating in terms of the information in your dataset.

**ANSWER: **

Q5. Calculate the average of the human rights index across all points in the sample. What does this tell you about the prevalence of human rights protections? Make sure to explain what this value is communicating.

**ANSWER: **

Q6. Produce a simple scatterplot with average income (GDP per capita) on the horizontal axis and the human rights protections (Human Rights Index) on the vertical axis.

Q7. Make the plot again, but this time add a trend line (also known as a line of best fit or a regression line) using the `abline()` command.

Q8. What does this line tell you about the relationship between development and human rights?

**ANSWER: **

Q9. Is this a causal relationship? Explain why or why not.

**ANSWER: **

Q10. Did you collaborate with anyone on this problem set? If so, list them here.

**ANSWER: **

Q11. Did you use generative AI on any part of this problem set? If so, identify which model you used and how you used it -- specifically, identify all questions you used it for and explain how you used it for each question.

**ANSWER: **
