---
title: "POLS 15 -- Section 5 Example Solution"
author: "Hunter C. Smith"
date: "`r Sys.Date()`"
output: html_document
---

We won't be going over any new code this week. Instead, this is a chance for you to practice the analytical tools we've learned so far in the course to prepare for next week's midterm exam. 

### Choose a Dataset

- `wb.RData`: Data from the World Bank for economic and environmental variables from 2005.
- `voters.RData`: Data from a 2016 USC / LA Times Presidential Poll of voter demographics and attitudes toward presidential candidates.
- `nba.RData`: Statistics for NBA teams from 2012-2016.

### Setup

Clear your environment, set your working directory, and upload your dataset of choice.

```{r}
rm(list = ls())
#setwd("/home/jovyan/UCSB_PS15/Section 5/")
load("voters.RData")
```

### Explore Your Dataset

Identify your sample size, number of variables, and variable names.

```{r}
nrow(usc4)
ncol(usc4)
names(usc4)
```

Before moving on, hypothesize a relationship between two variables in your dataset. Which is your IV, and which is your DV?

**IV: Education (`education`)** \
**DV: Probability of voting for Hillary Clinton (`clint_vote`)** \
**Hypothesis: education is positively correlated with probability of voting for Clinton**

### Univariate Analysis

For each variable, compute the mean, median, range, and standard deviation. What do these values tell you about the distribution of each variable?

```{r}
summary(usc4$education)
sd(usc4$education)

summary(usc4$clint_vote)
sd(usc4$clint_vote)
```

**Education: Mean $\approx$ median, so probably normally distributed. Small SD, so probably tight distribution.** \
**Clinton Vote: Mean > median, so positively skewed. Large SD, so lots of variation in values (probably lots of extreme values).** \

Now, create a distribution plot for each variable (with appropriate labels). Describe each of the distributions.

```{r}
hist(usc4$education,
     xlab = "Education",
     main = "Distribution of Education",
     col = "blue")

hist(usc4$clint_vote,
     xlab = "Prob. Clinton Vote",
     main = "Distribution of Prob. Clinton Vote",
     col = "red")
```

**Both distributions are bimodal. Modal responses for `clint_vote` are at extreme values while modal responses for `education` are near the center of the distribution.**

### Bivariate Analysis

Compute the covariance and correlation between your variables and interpret each -- what do they tell you about the relationship?

```{r}
cov(usc4$education, usc4$clint_vote)
cor(usc4$education, usc4$clint_vote)
```

**Covariance: There is a positive association between `education` and `clint_vote`.** \
**Correlation: There is a weak positive correlation between `education` and `clint_vote`.**

Now make a scatterplot of these variables. Is there a clear pattern in the data?

```{r}
plot(usc4$education, usc4$clint_vote,
     xlab = "Education",
     ylab = "Prob. Clinton Vote",
     main = "Scatterplot of Education and Prob. Clinton Vote")
```

**Difficult to tell visually, but correlation tells us that there should be a slight upward trend.**

Write the regression model you'll estimate with these variables using appropriate mathematical notation.

Equation: $$clint\_vote_i = \beta_0 + \beta_1 education_i + \epsilon_i$$

Now estimate your regression and add the line to your scatterplot. What does the line suggest about the relationship?

```{r}
mod1 <- lm(clint_vote ~ education, data = usc4)

plot(usc4$education, usc4$clint_vote,
     xlab = "Education",
     ylab = "Prob. Clinton Vote",
     main = "Scatterplot of Education and Prob. Clinton Vote")
abline(mod1, col = "blue", lwd = 2)
```

**Confirms that there is a positive relationship because the regression line has a positive slope.**

Finally, pull up the model summary and do the following:

- Interpret $\hat{\beta_0}$.
- Interpret $\hat{\beta_1}$.
- Specify your null hypothesis using both mathematical notation and words.
- Specify your alternate hypothesis using both mathematical notation and words.
- Interpret the $p$-value for $\hat{\beta_1}$ and state your conclusions regarding significance and your hypotheses.
- Interpret $R^2$.

```{r}
summary(mod1)
```

**$\hat{\beta_0}$: When a person has not received any education, their predicted probability of voting for Clinton is 5.19 percentage points.** \
**$\hat{\beta_1}$: A one-year increase in education is associated with a 3.52 percentage point increase in the probability of voting for Clinton.**

**$H_0: \beta_1 = 0$. Education does *not* affect the probability of voting for Clinton.** \
**$H_A: \beta_1 \neq 0$. Education *does* affect the probability of voting for Clinton.**

**The $p$-value associated with $\hat{\beta_1}$ is 0.0000000454. This is less than the standard significance threshold ($\alpha = 0.05$), so we have sufficient evidence to reject the null hypothesis.**

**$R^2$ = 0.02844, indicating that the specified model explains 2.84 percent of the variation we observe in the probability of voting for Clinton. Because this is a bivariate model, we can also say that accounting for education explains about 2.84 percent of the variation we observe in the probability of voting for Clinton.**

### Bonus -- Heterogeneous Effects

If you have time, choose another variable from your dataset that might influence your results (think back to the example from Week 3 with the terror data). Create a meaningful subset of your data and compare the means of your variables. Does it seem like there may be heterogeneous effects?

```{r}
mean(usc4$education[usc4$age >= mean(usc4$age)])
mean(usc4$education[usc4$age < mean(usc4$age)])

mean(usc4$clint_vote[usc4$age >= mean(usc4$age)])
mean(usc4$clint_vote[usc4$age < mean(usc4$age)])

usc4$age_bin <- ifelse(usc4$age >= mean(usc4$age), 1, 0) # creating a binary variable based on the age breakpoint used above
t.test(usc4$clint_vote ~ usc4$age_bin) # testing whether the difference in clint_vote based on age_bin is significant
```

**Very slight difference-in-means for `education` based on age, but larger for `clint_vote`. The $p$-value reported in the t-test demonstrates that this difference is not statistically significant, though, so there are likely not heterogeneous effects based on age.**