---
title: "Final Exam"
author: "Prof. Mahdavi"
date: "Spring 2022"
output: html_document
---

TIME LIMITS:
YOU HAVE 3 HOURS TO COMPLETE THIS EXAM.

YOU MUST SUBMIT YOUR CODE AND HTML FILE ON GAUCHOSPACE BY 3:00PM THURSDAY JUNE 9TH.

**If you cannot knit your code for any reason, you should still submit your code (.Rmd) file to Gauchospace. Submissions without a knitted HTML file will incur a 2-point penalty.**

RULES:
YOU MAY USE THE INTERNET, AND ANY MATERIAL FROM THE COURSE.
NO EMAIL, PHONE, OR COMMUNICATION WITH ANYONE ABOUT THE FINAL DURING THE 3-HOUR WINDOW.
FAILURE TO FOLLOW THIS RULE LEADS TO A ZERO IN THE COURSE.

PLEASE WRITE YOUR CODE AND YOUR TEXT ANSWER BELOW EACH QUESTION IN THE CODE CHUNK AND THE SPACE MARKED *Answer:* 

### Part 1. Short answer questions (12 pts total)

Q1. You toss a fair coin (i.e., one that gives you heads about half the time) 1000 times and record the outcome for each toss. If you were to graph a histogram or density plot of all the outcomes, would you roughly see a normal distribution? Explain why or why not in no more than two sentences. **(4 pts. total)**

*Answer:* 


Q2. True or False: In the model $Y = B_0 + B_1 X_1 + B_2 X_2 + \epsilon$, our estimate of $B_1$ will be biased if we omit a variable $X_3$ that is correlated with $X_1$ and $X_2$ but does not have a relationship with $Y$. Please answer whether this statement is true or false, and explain why in no more than two sentences. **(4 pts. total)**

*Answer:* 


Q3. We read an article this quarter by Regina Bateson. (3-4 sentences, 100 words max) **(4 pts. total)**
What is her research question and what is her dependent variable? *(1 pt.)*
What does she find? *(1 pt.)*
Are her results robust to alternative model specifications? *(1 pt.)*
Why or why not? *(1 pt.)*

*Answer:* 


### Part 2. Do property rights undermine development? (38 pts. total)

Political economy theory views property rights as the foundation of economic and political development. But do property rights that are too strong encourage opportunistic behavior? To answer this question, political scientist Alisha Holland designed a survey experiment in Colombia. Her study randomly assigned individuals to treatment groups with different property rights frames and asked them about their preferences for highway and construction projects funded by the state. The key treatment is whether or not a respondent saw the following message: "The state protects the right to private property because property owners invest in their homes and neighborhoods." You will be replicating one aspect of her study, which shows how property rights affect attitudes towards compensation for state-led highway projects that go through an individual's land.

Load the dataset, `survey.RData`. Here is a brief description of the variables:

* responseid - unique survey respondent ID number
* proprights - dummy variable for whether individual received property rights treatment (0 = no, 1 = yes)
* compensation - amount in thousands of pesos that respondent desires from government in compensation for highway going through their property
* female - dummy variable for whether respondent is female or not (0 = no, 1 = yes)
* age - age of respondent in five ordered groups (1 = 18-24, ..., 5 = 55 or older)
* education - dummy variable for whether respondent finished high school or not (0 = no, 1 = yes)
* wealth - wealth index of respondent in deciles (1 = poorest, ..., 10 = richest)

Q4. Set your working directory and load the dataset `survey.RData`. **(2 pts. total)**

```{r}

```


Q5. It's always good practice to begin by looking at and summarizing your data. **(4 pts. total)** 
What is your sample size? *(1 pt.)*
What is the mean value of `compensation`? *(1 pt.)*
What is the variance of this variable? *(1 pt.)*
What is the unit of analysis? *(1 pt.)*

```{r}

```

*Answer:* 


Q6. Now let's look at the distribution of `proprights`. **(6 pts. total)**
Plot the histogram of this variable. *(1 pt.)*
Interpret the histogram plot - how could you describe this distribution? *(2 pt.)*
How would you describe the *distribution of the mean* of this distribution? *(3 pt.)*

```{r}

```

*Answer:* 

Q7. Now let's examine the association between `proprights` and `compensation`. **(4 pts. total)**
What is the covariance of `proprights` and `compensation`? *(2 pt.)*
What is the correlation? *(2 pt.)*

```{r}

```


Q8. You want to know whether the property rights treatment affects compensation, controlling for age, gender, and wealth. **(4 pts. total)**
In this case, what is your main independent variable? *(1 pt.)* 
What is your dependent variable? *(1 pt.)*
State your null hypothesis. *(1 pt.)* 
State your alternative hypothesis. *(1 pt.)*

*Answer:* 


Q9. Plot the relationship between `wealth` (X) and `compensation` (Y) using a scatterplot. For full credit, you must label all axes in your graph, including the title. **(3 pts. total)**

```{r}

```



Q10. Run a regression that shows the relationship between the property rights treatment (X) and compensation (Y), controlling for age, female, and wealth. **(4 pts. total)**
Show the output of the regression. *(1 pt.)*
Substantively interpret the coefficient that shows the relationship between property rights and compensation. *(2 pt.)*
Interpret your p-value. *(1 pt.)*

```{r}

```

*Answer:* 


Q11. Intepret your findings in terms of your hypotheses. Did you find evidence for or against the null hypothesis? **(2 pts. total)**

*Answer:* 


Q12. A critic challenges the findings of the study by claiming that the author did not account for potential imbalance between people who went completed high school and those who did not. To evaluate their criticism, add the education dummy variable as an additional control to your model from Q10 and run a new regression. **(5 pts. total)**
Show the output of this regression. *(1 pt.)*
Interpret the coefficient on the education dummy variable and evaluate its statistical significance. *(2 pt.)*
Are the model results from Q10 robust to this change? *(2 pt.)*

```{r}

```

*Answer:* 


Q13. Can you say that we have identified a causal relationship between property rights and compensation? Why or why not? **(4 pts. total)**

*Answer:* 


Q14. **[BONUS QUESTION]** What is the most interesting and useful idea related to research methods that you learned in this class, which you will remember in the future? *(1 pt.)*

*Answer:* 

