---
title: "PS 15 Section 5 - Bringing it all together"
---

##REMINDERS
1. KNIT AS YOU GO!
2. ALWAYS SET YOUR WORKING DIRECTORY FIRST
3. MAKE SURE YOU KNOW WHAT YOUR VARIABLES ARE (X and Y, ORDER MATTERS)
4. KEEP CHUNKS TOGETHER EVEN IF THE PROBLEM SET SAYS DIFFERENTLY
5. PUT TEXTBOOK ANSWERS UNDER THE QUESTION
6. YOU CAN WORK TOGETHER, BUT PLEASE WRITE ANSWERS IN YOUR OWN WORDS (object names too)


To solidify your understanding of the material and to give you time to study for next week's midterm exam, this week we will not be learning any new code. Instead, the steps below will guide you through a basic statistical analysis of a relationship between two variables of your choosing using R code from the previous four sections. You already know this stuff. There will be no code provided. To answer the questions below reference the code you have used from previous sections and problem sets. 

You can take as much time as you want and work on this outside of class and you will not turn it in.

#### 1. Choose a dataset

There are three datasets to choose from. One from IR, one from AP, and one for the sports fans. Pick one that fits your interests. Below is a short description of each dataset. 
*These are not representative and shouldn't be used for "real research."*

 * wb -  International Relations - World Bank Data: Includes basic data on the economy and environment in the year 2005.  
 
 *  voters - American Politics - USC LA/Times Presidential Poll data from Oct 18, 2016 of voter's demographics and attitudes toward the presidential candidates
 
 *  nba - Sports - NBA Team Results from 2012 through 2016. 37 variables so you'll need the attached codebook. 
 


STEP 1. GET AND SET WORKING DIRECTORY
```{r}
getwd()
setwd("/home/jovyan/UCSB_PS15/Section 5")
```

STEP 2. LOAD YOUR DATASET
```{r}

```

STEP 3. How many variables do you have? What are their names? What is your sample size?

```{r}

```

STEP 4. Look at the codebook for your dataset. Using your knowledge of the subject, hypothesize a relationship between two of the variables. Which variable is your dependent variable? Independent variable?

STEP 5. Find the mean, median, range, standard deviation, and variance for your independent variable and dependent variable. 
```{r}

```

STEP 6. Look at the sampling distribution(histogram) of both your variables. 

```{r}

```

How would you describe the distribution?

STEP 7. Create a frequency distribution for both of your variables.

```{r}

```

STEP 8. Calculate the correlation and covariance between your independent and dependent variables. 

```{r}

```

What do they tell you?

STEP 9. Create a scatterplot with your independent variable on the X axis and your dependent variable on the Y axis. Add a regression line to your plot. 

```{r}

```

What does the relationship look like? 

Bonus: Do you think you have a heteroskedacity problem?

STEP 10. Write down the model of your regression. 


STEP 11. Show the summary of your bivariate regression. Interpret the intercept and the coefficient for your indepedent variable. A one-unit change in your indepedent variable is associated with what change in your dependent variable?

```{r}

```

Interpret your regression. 

* What are the intercept and the coefficient for your indepedent variable telling you?. 

* A one-unit change in your indepedent variable is associated with what change in your dependent variable?

* Is the relationship significant? ie interpret the p-value. 

* What is your null hypothesis? Do you reject it?

* What is your alternative hypothesis? Does the data support it?

STEP 12. Create a meaningful subset of your data. Run the same regression and see how/if results change. 



Bonus Activities:

* Choose a different independent and/or dependent variable (maybe a different dataset)

* Bedazzle (or at least properly label) your graphs.

* Create a boxplot for your independent variable

* Compare means of subsets (i.e. males/females, rich/poor, your team vs. a rival)
