---
title: "POLS 15 -- Section 5"
author: "Hunter C. Smith"
date: "`r Sys.Date()`"
output: html_document
---

We won't be going over any new code this week. Instead, this is a chance for you to practice the analytical tools we've learned so far in the course to prepare for next week's midterm exam. 

### Choose a Dataset

- `wb.RData`: Data from the World Bank for economic and environmental variables from 2005.
- `voters.RData`: Data from a 2016 USC / LA Times Presidential Poll of voter demographics and attitudes toward presidential candidates.
- `nba.RData`: Statistics for NBA teams from 2012-2016.

### Setup

Clear your environment, set your working directory, and upload your dataset of choice.

```{r}

```

### Explore Your Dataset

Identify your sample size, number of variables, and variable names.

```{r}

```

Before moving on, hypothesize a relationship between two variables in your dataset. Which is your IV, and which is your DV?

### Univariate Analysis

For each variable, compute the mean, median, range, and standard deviation. What do these values tell you about the distribution of each variable?

```{r}

```

Now, create a distribution plot for each variable (with appropriate labels). Describe each of the distributions.

```{r}

```

### Bivariate Analysis

Compute the covariance and correlation between your variables and interpret each -- what do they tell you about the relationship?

```{r}

```

Now make a scatterplot of these variables. Is there a clear pattern in the data?

```{r}

```

Write the regression model you'll estimate with these variables using appropriate mathematical notation.

Equation:

Now estimate your regression and add the line to your scatterplot. What does the line suggest about the relationship?

```{r}

```

Finally, pull up the model summary and do the following:

- Interpret $\hat{\beta_0}$.
- Interpret $\hat{\beta_1}$.
- Specify your null hypothesis using both mathematical notation and words.
- Specify your alternate hypothesis using both mathematical notation and words.
- Interpret the $p$-value for $\hat{\beta_1}$ and state your conclusions regarding significance and your hypotheses.
- Interpret $R^2$.

```{r}

```

### Bonus -- Heterogeneous Effects

If you have time, choose another variable from your dataset that might influence your results (think back to the example from Week 3 with the terror data). Create a meaningful subset of your data and compare the means of your variables. Does it seem like there may be heterogeneous effects? Try running a regression model to examine this.

```{r}

```

