---
title: "Problem Set 3"
author: "Prof. Paasha Mahdavi, PS 15, UCSB"
output: html_document
---

# Reminders about submitting your assignment
Please make sure that you first rename and save this file as "LASTNAME-pset3.Rmd" before you do any work. Once you are finished with your problem set and have knitted your file, please remember the following steps to properly upload your assignment to Canvas:

1. In the Files pane of RStudio, click the box beside all the files you want to download (i.e., the .Rmd and .html files).
2. Hit the gear icon and select export from the dropdown menu.
3. Hit download in the dialog box.
4. Find the zip folder in your downloads (NOT through your browser) and unzip the folder so you can access the individual .html and .Rmd files.
5. Go to the Canvas submission portal and submit the .Rmd and .html files contained within the unzipped folder.

As a reminder, you need to submit both your .html and .Rmd files for full credit. Any other file extension, such as .htm or .mhtml, cannot be read and graded. Also, please remember that you need to submit both of your individual files, NOT the zip folder. If you submitted the zip folder, please follow the steps above to correctly submit your problem set.

*Submit your HTML output and .Rmd file to Canvas by the deadline.*


***

## Question 1. What drives greenhouse gas emissions? 

One of the most important questions in environmental politics research is about the relationship between greenhouse gas emissions and economic development. In this problem set, we will study the determinants of emissions as a function of economic development, energy usage, and the carbon intensity of energy demand. To do so, we will use data from the Our World in Data project for countries in 2021. 

Variable Name  | Description 
------------- | ------------- 
country | Country - Geographic location.

iso_code | ISO code - ISO 3166-1 alpha-3 three-letter country codes.

greenhouse_gas_emissions | Emissions from electricity generation - Measured in megatonnes of CO2 equivalents.

primary_energy_consumption | Primary energy consumption - Measured in terawatt-hours.

electricity_demand | Electricity demand - Measured in terawatt-hours.

carbon_intensity_elec | Carbon intensity of electricity generation - Greenhouse gases emitted per unit of generated electricity, measured in grams of CO2 equivalents per kilowatt-hour.

gdp_cap | Gross domestic product (GDP) per capita

(a) Load the data set (energy.RData) using the `load()` function. 

```{r}

```

(b) It is always good practice to get a sense of the distributions of your main variables. Calculate the mean and standard deviation for *greenhouse_gas_emissions*, *gdp_cap*, and *carbon_intensity_elec*. What is the average amount of a country's level of emissions?

```{r}

```

*Answer 1b*

(c) We will start by testing the hypothesis that richer countries have higher emissions. For this, regress *greenhouse_gas_emissions* on *gdp_cap*. Show the `summary()` for a bivariate regression model and interpret the results, specifically the sample estimates  $\hat{\beta_1}$ $\hat{\beta_0}$ and their respective p-values. State your null and alternative hypotheses for this model. 

```{r}


```

*Answer 1c*

(d) Should the result in 1c be given a causal interpretation? Why or why not? Describe one possible confounder. 

*Answer 1d*

(e) Now we will use multivariate regression to begin addressing some of the possible confounders. In other words, we will build a more complex model. 

Next let's evaluate whether the previous finding is robust to the accounting for the energy usage of a country, measured with *primary_energy_consumption*, and the carbon-intensity of electricity demand, measured with *carbon_intensity_elec*. The latter is an indication of whether a country's electricity is generated with more carbon-intense fuels like coal versus non-carbon-intensive fuels like solar, wind, nuclear, and hydro. 

Run a multivariate OLS model that regresses emissions on country wealth, energy usage, and carbon-intensity of energy usage. That is, regress  *greenhouse_gas_emissions* on  *gdp_cap*, but controlling for  *primary_energy_consumption* and *carbon_intensity_elec* by including all three variables in the model.

Produce a summary table for your model results and correctly interpret these results. That is:

- Explain the meaning of all the coefficient estimates

- Remark on the significance of each coefficient (except the intercept)

- Most importantly, compared to the previous model, remark on any changes to the estimated coefficient for *gdp_cap* and its statistical significance. What is going on here?

```{r}

```

*Answer 1e*

(f) Can you make a causal interpretation for both energy usage and carbon intensity? What other potential confounders remain? 

*Answer 1f*



## Question 2. Substantive Interpretation 

For question 1, explain briefly (300 words or less) what did you learn from analyzing the predictors of greenhouse gas emissions. Your answer must reference specific results (i.e. magnitude and statistical significance of at least one coefficient). Please make sure to correctly interpret these findings in the context of a multivariate regression. To get full points, your response should (1) briefly describe one take-away from the analysis, (2) discuss the magnitude and statistical significance of that coefficient, and (3) explain these in relation to the other variables in the model

*Answer 2*

## Question 3 (no points, ungraded).

Did you use generative AI to complete this problem set? If so, what software or type of AI did you use (e.g., chatGPT, Gemini, etc.)? Remember, there is no penalty for doing so, as long as you disclose your usage of AI. 

*Answer 3*