---
title: "PS 15 Section 9 - Data Management"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today, we will be looking at a data set by Golder and Clark (2006) to make predictions about the number of parties in a political system. The dataset is not as orderly and clean as we need it to be for analysis. Therefore, we will explore the data and prepare it.

1. Begin by setting your working directory and importing the data file.

```{r}
#setwd("")
load("clarkgolder1.RData")
```

2. Let's have a look at the variables contained in the dataset:

```{r}
str(cg1)
```

The data set contains 5 variables:
country = the country where the election took place
year = the year of the election
avemag = average district magnitude, the number of members to be elected in each electoral district
eneg = effective number of ethnic groups
enep = effective number of parties

Take a look at the data types of the variables. Some variables that should be saved as vector of type numeric or double (R's way of saying "decimal") are actually saved as character vectors.

Moreover, some value in the eneg and enep variable are ".". This is one way of saving NAs but R does not register them as NA. Let's tackle these problems one by one.

3. First, let's replace the "." by "NA". The easiest way to do this is to use the ifelse function, which you are familiar with from creating dummy variables. We tell R to check whether our vector contains "." and if it does, we tell it to replace it with NA. If it does not, it should take the original value.

```{r}
cg1$avemag <- ifelse(cg1$avemag==".", NA, cg1$avemag)
cg1$eneg <- ifelse(cg1$eneg==".", NA, cg1$eneg)
cg1$enep <- ifelse(cg1$enep==".", NA, cg1$enep)
```

4. The second problem we are facing is that vectors are of type character when they should be numeric or double. Luckily there is an easy fix. We use the function as.numeric and as.double to convert the vectors.

```{r}
cg1$avemag <- as.numeric(cg1$avemag)
cg1$eneg <- as.double(cg1$eneg)
cg1$enep <- as.double(cg1$enep)

str(cg1)
```

5. Let's run a multivariate regression to tease out the impact of electoral institutions and sociological factors on the number of parties. We will use "avemag"" and "eneg". "Eneg"" will be our proxy for social heterogeneity.  
```{r}
model1 <- lm(enep~avemag + eneg, data=cg1)
summary(model1)
```


6. For our next analysis we need a variable that is in a different data set. To consolidate them into one, we will merge them. To do that, R needs to know how to align the rows so it adds the new values to the right cases. In our data set, the rows are uniquely identified by a country name and a year. This is what we will provide to the merge function.

First we load the second data set.

```{r}
load("clarkgolder2.RData")
```

Then we merge the two using the merge function. We will save the output to a new data frame.

```{r}
cg3 <- merge(cg1, cg2, c("country", "year"))
```

7. With the new data frame, let's run another regression model. This time, we'll add the "newdem" variable. The authors wanted to test whether or not electoral institutions still have an impact on the number of parties in newer democracies. Here new democracies refer to those democracies that formed after 1989. What kind of variable is "newdem"?

```{r}
model2 <- lm(enep~avemag + eneg + newdem, data=cg3)
summary(model2)
```

