---
title: "POLS 15 -- Section 9"
author: "Hunter C. Smith"
date: "`r Sys.Date()`"
output: html_document
---

### DIY Setup

As always, clear your environment, set your working directory, and load in today's first data file, `clarkgolder1.RData`.

```{r}
# Type your code here!
```

Here's a description of the variables we'll be using today:

Variable Name  | Description 
-------------  | ------------------------------------------------------------------ 
`country`        | country where an election took place
`year`           | year of the election
`avemag`         | average district magnitude, the number of members to be elected in each electoral district
`eneg`           | effective number of ethnic groups
`enep`           | effective number of political parties

### Data Processing

Many times, the data we work with isn't as "clean" as we'd like it to be, and today's dataset is a prime example. Let's look at a quick summary of the dataset using `str()`.

```{r}
str(cg1)
```

Right off the bat, there are two problems. First, some variables that are obviously numeric are considered character variables (what does this mean?). Second, we have some missing values coded as `.`, which `R` doesn't recognize, so we need to correct them.

Let's start by recoding the missing values currently denoted by `.` to `NA`. We can do this using `ifelse()`, with which you should be familiar at this point.

```{r}
cg1$avemag <- ifelse(cg1$avemag == ".", NA, cg1$avemag)
cg1$eneg <- ifelse(cg1$eneg == ".", NA, cg1$eneg)
cg1$enep <- ifelse(cg1$enep == ".", NA, cg1$enep)
```

Double-check that these changes worked as intended:

```{r}
str(cg1)
```

Next, let's change the data type of the numeric variables that are currently coded as character variables.

```{r}
cg1$avemag <- as.numeric(cg1$avemag)
cg1$eneg <- as.double(cg1$eneg)
cg1$enep <- as.double(cg1$enep)
```

Again, double-check that the changes were successful:

```{r}
str(cg1)
```

### DIY -- Multivariate Regression Interpretation

Next, we're going to practice multivariate regression interpretation. Use the code chunk below to regress `enep` on `avemag` and `eneg`, and interpret $\hat{\beta_1}$ and $\hat{\beta_2}$.

```{r}
# Type your code here!
```

$\hat{\beta_1}$ Interpretation:

$\hat{\beta_2}$ Interpretation:

### Data Merging

There is another variable called `newdem` which indicates whether a democracy was formed after 1989, which the authors consider "new". It should probably be included in our model, but it's saved in a different dataset. To include it, we need to merge the datasets together. `R` can do this, but first it needs to know at least one unique identifier so it can combine the datasets correctly. Let's load in the second dataset and see what identifier(s) are shared between the datasets.

```{r}
load("clarkgolder2.RData")
str(cg2)
```

As it turns out, we can merge the dataset according to country-years because both `country` and `year` are common across both datasets. Now, we can use `merge()` to combine the datasets.

```{r}
cg3 <- merge(cg1, # The name of the first dataset
             cg2, # The name of the second dataset
             c("country", "year") # The unique identifiers used to match observations
             )
```

Now, add `newdem` as an independent variable to the model and interpret $\hat{\beta_1}$, $\hat{\beta_2}$, and $\hat{\beta_3}$. Be sure to evaluate the robustness of $\hat{\beta_1}$ and $\hat{\beta_2}$ relative to the previous multivariate model.

```{r}
# Type your code here!
```

$\hat{\beta_1}$ Interpretation:

$\hat{\beta_2}$ Interpretation:

$\hat{\beta_3}$ Interpretation:

### Knit to HTML

Always remember to knit your file when you finish anything!