---
title: "Midterm Exam"
author: "Prof. Mahdavi"
date: "Fall 2022"
output: html_document
---

TIME LIMITS:
YOU HAVE 75 MINUTES TO COMPLETE THIS EXAM, STARTING FROM 11:00AM UNTIL 12:15PM.

YOU MUST SUBMIT YOUR CODE AND HTML FILE ON GAUCHOSPACE BY 12:15PM PACIFIC TIME ON THURSDAY NOVEMBER 3rd.

**If you cannot knit your code for any reason, you should still submit your code (.Rmd) file to Gauchospace. Submissions without a knitted HTML file will incur a 2-point penalty.**

RULES:
YOU MAY USE THE INTERNET AND ANY MATERIAL FROM THE COURSE.
NO EMAIL, PHONE, OR COMMUNICATION WITH ANYONE ABOUT THE MIDTERM DURING THE 75-MINUTE WINDOW.
FAILURE TO FOLLOW THIS RULE LEADS TO A ZERO IN THE COURSE.

PLEASE WRITE YOUR CODE AND YOUR TEXT ANSWER BELOW EACH QUESTION IN THE CODE CHUNK AND THE SPACE MARKED *Answer:* 

Q1, Q2, AND Q7 ONLY INVOLVE CODE, SO WRITE YOUR ANSWER INSIDE THE PROVIDED CODE CHUNK

### Midterm elections

What explains congressional midterm election outcomes? Political scientists have developed numerous models to explain the vote share of congressional candidates in midterm elections, with the simplest model being a function of the district's electoral support for the candidate's party in the most recent presidential election. 

Anthony Fowler and Ethan Bueno de Mesquita assembled a dataset to answer this question for the 2010 midterms based on district-level results from the 2008 presidential election. They examine the hypothesis that higher vote shares for the candidate's party in the 2008 election correspond with higher vote shares for the candidate in the 2010 midterm election. 

Download the dataset, `congress.RData`, which you'll find in the Midterm working directory of the RStudio online platform. Here is a brief description of the variables:

* state_dist - label indicating the state and congressional district.
* representative - member of Congress that represented this district in this term. You may not need to use this variable in your analysis, but you at least have it as a reference.
* republican - takes a value of 1 if the representative was a Republican and 0 if the representative was a Democrat. 
* presvote2008 - average Republican two-party vote share in that particular congressional district from the 2008 presidential election.
* congvote2010 - the Republican two-party vote share in the 2010 congressional election. For example, a value of .51 means that a Republican candidate won by 1 percentage point and went on to serve in congress in 2011. A value of .49 would indicate that a Democrat won by 1 percentage point. 


Q1. Set your working directory and load the dataset `congress.RData` (and make sure you are capitalizing both the `R` and the `D` in `.RData`). **(1 pt total)**

```{r}

```


Q2. It's always good practice to begin by looking at and summarizing your data. **(2 pts total)** 

a. What is your sample size? *(0.5 pt.)*
b. What is the mean value of `presvote2008`? *(0.5 pt.)*
c. What is the median value of `presvote2008`? *(0.5 pt.)*
d. What is the variance of `presvote2008`? *(0.5 pt.)*

```{r}

```


Q3. Let's think about variance in more detail. **(3 pts total)**

a. What is the mathematical formula for population variance, using the notation of expectations (e.g., $E[X]$)? *(1 pt.)*
b. How would you describe the concept of variance in one sentence? *(1 pt.)*
c. What happens to the sample variance of a random variable if the sample size increases, holding everything else fixed? *(1 pt.)*

*Q3 Answer: *


Q4. Now let's look at the distribution of `presvote2008`. **(2 pts total)**

a. Plot the distribution using a histogram. *(1 pt.)*
b. Interpret the histogram - how could you describe this distribution in words? *(1 pt.)*

```{r}

```

*Q4 Answer: *


Q5. Now let's examine the association between `presvote2008` and `congvote2010`. **(2 pt total)**

a. What is the covariance between `presvote2008` and `congvote2010`? *(0.5 pt.)*
b. What is the correlation? *(0.5 pt.)*
c. How are covariance and correlation related? *(1 pt.)*


```{r}

```

*Q5c Answer:*

Q6. You want to know whether past presidential election results predict current congressional election results. **(3 pts total)**

a. In this case, what is your independent variable? *(0.5 pt.)* 
b. What is your dependent variable? *(0.5 pt.)*
c. State your null hypothesis. *(1 pt.)* 
d. State your alternative hypothesis. *(1 pt.)*

*Q6 Answer: *


Q7. Plot the relationship between `presvote2008` and `congvote2010` using a scatterplot. For full credit, you must label all axes in your graph, including the title. **(1 pt total)**

```{r}

```


Q8. Run a regression of past presidential election results on current congressional election results. **(3 pts total)**

a. Show the output of the regression. *(1 pt.)*
b. Substantively interpret the slope coefficient. *(1 pt.)*
c. Interpret your p-value. *(1 pt.)*

```{r}

```

*Q8 Answer: *


Q9. Interpret your findings in terms of your hypotheses. Did you find evidence for or against the null hypothesis? **(2 pts total)**

*Q9 Answer: *


Q10. Can you say that we have identified a causal relationship between past presidential election results and current congressional election results? Why or why not? **(3 pts total)**

*Q10 Answer: *


Q11. We read an article this week by Kalla \& Broockman. (4 sentences, 100 words max) **(3 pts total)**

a. What is their research question? *(0.5 pt.)*
b. What do they find? *(1 pt.)*
c. Are they able to infer a causal effect? *(0.5 pt.)*
d. Why or why not? *(1 pt.)*

*Q11 Answer: *
