---
title: "Problem Set 3 -- Solution Set"
author: "Hunter C. Smith"
date: "Summer 2026"
output: html_document
---

## Question 1. What drives greenhouse gas emissions? 

(a) Load the dataset (`electricity_data.RData`) into your environment (there are no points associated with this). Then, calculate the mean, median, and standard deviation for `greenhouse_gas_emissions` and `gdp_cap`. For each variable:

- Interpret the mean value using appropriate units.
- Compare the mean and median. What does this tell you about the direction of skewness for each variable? Do NOT use a distribution plot -- rely only on your summary statistics.

**(2 points)**

```{r}
rm(list = ls())
setwd("/Users/hunte/Desktop/POL SW 15 Materials (M26)/PSets")
load("electricity_data.RData")

mean(energy$greenhouse_gas_emissions)
median(energy$greenhouse_gas_emissions)
sd(energy$greenhouse_gas_emissions)

mean(energy$gdp_cap)
median(energy$gdp_cap)
sd(energy$gdp_cap)
```

**ANSWER: The average country's GHG emissions in this dataset is 83.51 Mt CO$_2$e. The average country's GDP per capita in this dataset is \$18,805.94. For GHG emissions, mean \> median by a large margin, indicating a positive skew. For GDP per capita, mean \> median, indicating a positive skew.**

(b) We will start by testing the hypothesis that richer countries have higher emissions. Regress `greenhouse_gas_emissions` on `gdp_cap` and show the `summary()` output. Interpret the results, addressing all of the following:

- What are your independent and dependent variables?
- Report your estimates for $\beta_0$ and $\beta_1$ to 3 decimal places, and explain what each estimate is communicating using its specific value and appropriate units. 
- Report the $t$-statistic for $\hat{\beta_1}$ and interpret it against the critical value consistent with the standard significance threshold $\alpha$.
- Report the $p$-value for $\hat{\beta_1}$. Also specify the null and alternate hypotheses you are testing using appropriate mathematical notation. Finally, evaluate the p-value against the standard significance threshold $\alpha$ and provide a conclusion regarding your hypotheses.
- Do your $t$-statistic and $p$-value lead you to the same conclusions regarding statistical significance? Explain why or why not.
- Report the model's $R^2$ value to 3 decimal places and provide an interpretation -- what is this specific value communicating about your model?

**(3 points)**

```{r}
mod1 <- lm(greenhouse_gas_emissions ~ gdp_cap, data = energy)
summary(mod1)
```

**ANSWER:** \

**IV = GDP per capita, DV = GHG emissions** \

**$\hat{\beta_0}$ = 52.121. When a country's GDP per capita is \$0, its GHG emissions are predicted to be 52.121 Mt CO$_2$e.** \

**$\hat{\beta_1}$ = 0.002. A one-dollar increase in GDP per capita is associated with a 0.002 Mt CO$_2$e increase in GHG emissions.** \

**$t$ = 0.981. This is below $t_c$ = 1.96, so $\hat{\beta_1}$ is statistically insignificant and we fail to reject the null hypothesis.** \

**$p$ = 0.328. $H_0: \beta_1 = 0$. $H_A: \beta_1 \neq 0$. The $p$-value is larger than $\alpha$ = 0.05, so $\hat{\beta_1}$ is statistically insignificant and we fail to reject the null hypothesis.** \

**The $p$-value and $t$-statistic for any given coefficient will always lead you to the same conclusions because the $p$-value is calculated based on where the $t$-statistic falls in the $t$-distribution. It represents the probability of getting a result as extreme as we did (i.e., a $t$-statistic as large as we estimated) if $H_0$ is true.** \

**$R^2$ = 0.006. Because this is a bivariate model, we can say that accounting for GDP per capita explains roughly 0.6% of the observed variation in GHG emissions.**

(c) Using your estimates for $\beta_0$ and $\beta_1$ from part (b), predict GHG emissions for Sweden and Finland using their observed GDP per capita values. You must use `R` code to retrieve each country's GDP from the dataset -- do NOT open the dataset directly. Show all calculations in your code chunk, and report each prediction in your written answer. Then, for each country, calculate the residual (again, use code to retrieve the relevant observed values). What do these residuals tell you about how well the bivariate model predicts GHG emissions for these two countries? **(2 points)**

```{r}
sweden_gdp <- energy$gdp_cap[energy$country == "Sweden"]
finland_gdp <- energy$gdp_cap[energy$country == "Finland"]

sweden_pv_biv <- 52.121367 + (0.001669 * sweden_gdp)
finland_pv_biv <- 52.121367 + (0.001669 * finland_gdp)

sweden_pv_biv
finland_pv_biv

sweden_resid <- sweden_pv_biv - energy$greenhouse_gas_emissions[energy$country == "Sweden"]
finland_resid <- finland_pv_biv - energy$greenhouse_gas_emissions[energy$country == "Finland"]

sweden_resid
finland_resid
```

**ANSWER:** \

**Sweden: Predicted GHG emissions = 128.89 Mt CO$_2$e. Residual = 121.72** \

**Finland: Predicted GHG emissions = 118.91 Mt CO$_2$e. Residual = 109.14** \

**The model isn't doing particularly well at predicting GHG emissions for these two countries given the large residuals, which indicate that the model is severely overestimating these values compared to the observed values.**

(d) Can the coefficient estimate for `gdp_cap` from (b) be interpreted as the causal effect of GDP per capita on GHG emissions? Thoroughly explain why or why not. **(1 point)**

**ANSWER: The coefficient estimate for `gdp_cap` cannot be interpreted as a causal effect. This is an observational study rather than an experiment, meaning that GDP per capita was not (can feasibly/ethically cannot) be randomly assigned across countries. This makes our estimate susceptible to omitted variable bias from confounding variables. Without accounting for all potential confounders, we cannot conclude a causal effect of GDP on GHG emissions.**

(e) Using values from part (b)'s output, calculate and report the upper and lower bound of the 95% confidence interval associated with your $\hat{\beta_1}$. Explain what this confidence interval is communicating about your analysis. Based on this confidence interval, is your $\hat{\beta_1}$ statistically significant? Explain why or why not. **(1 point)**

```{r}
upper <- 0.001669 + (1.96 * 0.001701)
lower <- 0.001669 - (1.96 * 0.001701)

upper
lower
```

**ANSWER: Upper bound = 0.00500296. Lower bound = -0.00166496. This means that, and a 95% confidence level, the range of possible $beta_1$ values that are consistent with our point estimate ranges from a minimum $beta_1$ of -0.0017 to a maximum $beta_1$ of 0.005. The confidence interval suggests that our $\hat{\beta_1}$ is statistically insignificant because the confidence interval contains zero.**

## Question 2. Confidence in Estimates

Two researchers are studying political news consumption among voters. Using two random samples, they find the following:

- Researcher A: $\overline{X}$ = 7 hours, $s^2$ = 5, $n$ = 300
- Researcher B: $\overline{X}$ = 7 hours, $s^2$ = 5, $n$ = 400

Based on what you know about $\mathbb{E}[\overline{X}]$ and $Var(\overline{X})$, which researcher should be more confident in their estimate of $\overline{X}$? Explain your logic. **(1 point)**

**ANSWER: Researcher B should be more confident because the sample variance is the same but researcher B has a larger sample size, meaning that the variance of the sample mean will be smaller for them, which increases precision.**

## Question 3.

Did you collaborate with anyone on this problem set? If so, list them here.

**ANSWER: -0.5 if left unanswered.**

## Question 4.

Did you use generative AI on any part of this problem set? If so, identify which model you used and how you used it -- specifically, identify all questions you used it for and explain how you used it for each question.

**ANSWER: -0.5 if left unanswered.**
